Interpolation of Sparse High-Dimensional Data\thanks {This work was supported by the National Science Foundation Grant CNS-1565314.}

### 2019-03-20 Paper_[2019-02]_Numerical_Algorithms/revision-one.tex
created an affine weights visual. finished NA paper draft. wrote report on Xiaodong clock cycle data.

% Insert here the call for the packages your document requires
\usepackage{graphicx}
\usepackage{tikz}     % For adding axes to plots
\usetikzlibrary{positioning}
\usepackage{graphicx} % For including PDFs
\usepackage{amsmath}  % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs} % For formal tables
\usepackage{tabularx} % For tables with nice column widths.
\usepackage[colorlinks,allcolors=blue]{hyperref}
\usepackage{float}    % For controlling the placement of tables.

@@

\newcolumntype{d}[1]{D{.}{.}{#1}}

@@

\begin{abstract}
Increases in the quantity of available data have allowed all fields of
science to generate more accurate models of multivariate phenomena.
Regression and interpolation become challenging when the dimension of
data is large, especially while maintaining tractable computational
complexity. Regression is a popular approach to solving approximation
problems with high dimension, however there are often advantages to
interpolation. This paper presents a novel and insightful error bound
for (piecewise) linear interpolation in arbitrary dimension and
contrasts the performance of some interpolation techniques with
popular regression techniques. Empirical results demonstrate the
viability of interpolation for moderately high dimensional
approximation problems, and encourage broader application of
interpolants to multivariate approximation in science.

@@

This section presents the theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straight forward extensions of this error
bound.

@@

In summary, the approximation error of a linear (simplicial)
interpolant tends quadratically towards zero when approaching observed
data only when the diameter of the simplex is also reduced at a
proportional rate.  Only linear convergence to the true function can
be achieved in practice, without the incorporation of additional
observations. Notice that the approximation error is largely
determined by the spacing of observed data. Predictions made by
simplices whose vertices are not well-spaced (i.e. have large
diameter, or are nearly contained in a hyperplane) have higher
error. In light of this error bound, an empirical evaluation of
presented algorithms follows.

%% ===================================================================
\vspace{-2mm}
\section{Data and Empirical Analysis}
\label{sec:data}

The theoretical results constructed in Section \ref{sec:theory} for
(piecewise) linear interpolation are promising and apply directly to
Delaunay interpolation, however they are difficult to interpret in
context with approximation algorithms that do not have similar known
uniform error bounds. For that reason, this paper utilizes five
different data sets of varying dimension and application to construct
approximations and compare the accuracy of different techniques.

@@

All approximations are constructed using $k$-fold cross validation as
described in \cite{kohavi1995study} with $k=10$. This approach
randomly partitions data into $k$ equal sized (nearly) sets. Each
algorithm is then evaluated by constructing an approximation over each
unique union of $k-1$ elements of the partition, making predictions
for points in the remaining element. As a result, each observed data
point is used in the construction of $k-1$ different approximations
and is approximated exactly once. The $k$-fold cross validation method
is data-efficient and provides an unbiased estimate of the expected
prediction error \cite{kohavi1995study}, however it should be noted
that neither this method nor others can provide a universally unbiased
estimator for the variance of prediction error \cite{bengio2004no}.

@@

\begin{table}
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l}
    Algorithm & \multicolumn{2}{c|}{Avg. \% Best} & \multicolumn{2}{c|}{\multilinecell{Avg. Fit or \\ Prep. Time (s)}} & \multicolumn{2}{c}{Avg. App. Time (s)}\\
    \hline
    MARS & \quad\quad$4$&$6$ & \,\quad$22$&$3$ & \quad\quad$0$&$00137$\\
    SVR & $\mathit{21}$&$\mathit{1}$ & $0$&$503$ & $\mathit{0}$&$\mathit{000137}$\\
    MLP & $\mathbf{42}$&$\mathbf{1}$ & $210$&$0$ & $0$&$00102$\\
    Delaunay & $6$&$0$ & $\mathbf{0}$&$\mathbf{000143}$ & $3$&$03$\\
    ShepMod & $18$&$4$ & $0$&$711$ & $0$&$000149$\\
    LSHEP & $8$&$4$ & $1$&$69$ & $\mathbf{0}$&$\mathbf{000116}$\\
    Voronoi & $0$&$5$ & $\mathit{0}$&$\mathit{127}$ & $0$&$0375$\\
    BoxSplines & $0$&$2$ & $0$&$816$ & $0$&$000526$\\
  \end{tabular}
  \caption{This average of Appendix Tables
    \ref{table:best-forest-fire}, \ref{table:best-parkinsons},
    \ref{table:best-weather}, and \ref{table:best-credit-card}
    provides a gross summary of overall results. The columns display
    (weighted equally by data set, \textit{not} points) the average
    frequency with which any algorithm provided the lowest absolute
    error approximation, the average time to fit / prepare, and the
    average time required to approximate one point. Interpolants
    provide the lowest error approximation one third of the time
    across all data, while regressors occupy the other two
    thirds. This result is obtained without any customized tuning or
    preprocessing to maximize the performance of any given
    algorithm. In practice, tuning and preprocessing may have large
    effects on approximation performance.}
  \label{table:avg-performance}
\end{table}

@@

The major advantages to interpolation lie in the near absence of
\textit{fit} time. Delaunay, LSHEP, and ShepMod all require pairwise
distance calculations, for numerical robustness (Delaunay) and to
determine the radii of influence for data points (LSHEP and
ShepMod). At least hundreds, and sometimes hundreds of thousands of
predictions can be made by the interpolants before the most widely
used regressor (MLP) finishes fitting these relatively small data
sets.  However, the computational complexities of all interpolants
presented are greater than linear in either dimension or number of
points, whereas the regressors' nonlinear complexity in dimension
generally comes from the model fitting optimization.

The new theoretical results presented in Section \ref{sec:theory}
directly apply to Delaunay interpolation, however the performance of
Delaunay does not appear significantly better than other algorithms on
these data sets.  This observation may be due to the stochastic nature
of the data, but it also speaks to the power of the approximations
generated by the different interpolation methods.  The strong
performance of other interpolants suggests that theoretical results
similar to those presented in this work can be achieved for the other
interpolants under reasonable assumptions.

Finally, most of the interpolants presented in this work benefit from
the ability to model any function over real $d$-tuples with a range
that is closed under convex combinations. The results of the
distribution prediction case study indicate that interpolants can
effectively predict CDFs. The error analysis for that work relies on
the KS statistic, which captures the worst part of any prediction and
hence provides a conservatively large estimate of approximation
error. The average absolute errors in the predicted CDFs are always
lower than the KS statistics. However, the KS statistic was chosen as
a metric because of the important surrounding statistical theory. A
nonnegligible volume of predictions provide impressively low levels of
average absolute error in that final case study.

@@

The major contributions of this work are: 1) new uniform theoretical
error bounds for piecewise linear interpolation in arbitrary dimension
(Section \ref{sec:theory}); and 2) an empirical evaluation across
real-world problems (Section \ref{sec:data}) that demonstrates
interpolants produce competitively accurate models of multivariate
phenomenon when compared with common regressors for sparse, moderately
high dimensional problems.

The various interpolants discussed in this paper have been
demonstrated to effectively approximate multivariate phenomena up to
$30$ dimensions. The underlying constructions are theoretically
straightforward, interpretable, and yield reasonably accurate
predictions. Most of the interpolants' computational complexities make
them particularly suitable for applications in even higher
dimension. The major benefits of interpolation are seen when only a
small number of approximations ($\leq 1000$) are made from data and
when there are relatively few data points for the dimension (for
empirical results presented, $\log_d n \leq 5$).

@@

\begin{appendix}
\section*{Appendix}
\label{sec:appendix}

\vspace*{\fill}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS & $0$&$00984$ & $3$&$11$ & $7$&$01$ & $\mathit{11}$&$\mathit{7}$ & $1090$&$0$\\
    SVR & $\mathit{0}$&$\mathit{00402}$ & $\mathbf{0}$&$\mathbf{243}$ & $\mathbf{0}$&$\mathbf{416}$ & $\mathbf{6}$&$\mathbf{19}$ & $1090$&$0$\\
    MLP & $0$&$0426$ & $2$&$63$ & $\mathit{5}$&$\mathit{27}$ & $14$&$0$ & $1090$&$0$\\
    Delaunay & $\mathbf{0}$&$\mathbf{00}$ & $1$&$98$ & $5$&$37$ & $13$&$1$ & $\mathit{1080}$&$\mathit{0}$\\
    ShepMod & $\mathbf{0}$&$\mathbf{00}$ & $\mathit{1}$&$\mathit{93}$ & $6$&$27$ & $16$&$0$ & $1090$&$0$\\
    LSHEP & $0$&$0400$ & $2$&$17$ & $8$&$87$ & $19$&$1$ & $\mathbf{1070}$&$\mathbf{0}$\\
    Voronoi & $0$&$00982$ & $3$&$65$ & $7$&$56$ & $15$&$6$ & $1090$&$0$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{00}$ & $2$&$27$ & $5$&$91$ & $16$&$4$ & $1090$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-forest-fire}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS & \quad$\mathit{7}$&$\mathit{7}$ &  & \quad\,\,\,$29$&$1$ & \quad$0$&$00137$ & \quad\quad\quad$0$&$0686$\\
    SVR & $\mathbf{80}$&$\mathbf{2}$ &  & $0$&$00584$ & $\mathit{0}$&$\mathit{0000620}$ & $\mathit{0}$&$\mathit{00310}$\\
    MLP & $0$&$0$ &  & $32$&$8$ & $0$&$000871$ & $0$&$0436$\\
    Delaunay & $3$&$5$ &  & $\mathbf{0}$&$\mathbf{0000249}$ & $0$&$0385$ & $1$&$93$\\
    ShepMod & $3$&$7$ &  & $0$&$00634$ & $0$&$0000644$ & $0$&$00322$\\
    LSHEP & $5$&$1$ &  & $0$&$0275$ & $\mathbf{0}$&$\mathbf{0000463}$ & $\mathbf{0}$&$\mathbf{00231}$\\
    Voronoi & $0$&$0$ &  & $\mathit{0}$&$\mathit{000182}$ & $0$&$000396$ & $0$&$0198$\\
    BoxSpline & $0$&$4$ &  & $0$&$00724$ & $0$&$0000978$ & $0$&$00489$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in Table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.}
  \label{table:best-forest-fire}
\end{table}

\vspace*{\fill}
\newpage

\begin{table}
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS & $0$&$00948$ & $9$&$98$ & $20$&$4$ & $32$&$9$ & $863$&$0$\\
    SVR & $0$&$00138$ & $3$&$17$ & $7$&$31$ & $12$&$6$ & $\mathbf{27}$&$\mathbf{5}$\\
    MLP & $0$&$0000239$ & $\mathit{0}$&$\mathit{533}$ & $\mathit{1}$&$\mathit{25}$ & $\mathbf{2}$&$\mathbf{84}$ & $39$&$3$\\
    Delaunay & $\mathit{3}$&$\mathit{72 \times 10^{-12}}$ & $1$&$20$ & $3$&$50$ & $7$&$67$ & $30$&$7$\\
    ShepMod & $\mathbf{0}$&$\mathbf{00}$ & $\mathbf{0}$&$\mathbf{255}$ & $\mathbf{0}$&$\mathbf{908}$ & $\mathit{3}$&$\mathit{43}$ & $34$&$5$\\
    LSHEP & $0$&$00254$ & $2$&$93$ & $7$&$16$ & $13$&$1$ & $\mathit{29}$&$\mathit{0}$\\
    Voronoi & $\mathbf{0}$&$\mathbf{00}$ & $1$&$29$ & $3$&$52$ & $6$&$87$ & $30$&$1$\\
    BoxSpline & $0$&$00287$ & $4$&$39$ & $9$&$10$ & $14$&$8$ & $41$&$3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS & \quad\,$0$&$0$ &  & \quad\quad$37$&$9$ & \quad\,\,$0$&$00253$ & \quad\quad\quad\,\,$1$&$48$\\
    SVR & $0$&$1$ &  & $0$&$859$ & $\mathit{0}$&$\mathit{000181}$ & $\mathit{0}$&$\mathit{106}$\\
    MLP & $\mathit{32}$&$\mathit{0}$ &  & $348$&$0$ & $0$&$00111$ & $0$&$653$\\
    Delaunay & $0$&$0$ &  & $\mathbf{0}$&$\mathbf{000190}$ & $3$&$69$ & $2170$&\\
    ShepMod & $\mathbf{66}$&$\mathbf{4}$ &  & $1$&$13$ & $0$&$000182$ & $0$&$107$\\
    LSHEP & $0$&$0$ &  & $2$&$39$ & $\mathbf{0}$&$\mathbf{000144}$ & $\mathbf{0}$&$\mathbf{0845}$\\
    Voronoi & $1$&$6$ &  & $\mathit{0}$&$\mathit{298}$ & $0$&$0274$ & $16$&$1$\\
    BoxSpline & $0$&$0$ &  & $1$&$26$ & $0$&$000643$ & $0$&$377$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Parkinson's data in Table
    \ref{table:error-parkinsons}. On the right columns are median fit
    time of 5288 points, median time for one approximation, and median
    time approximating 587 points.}
  \label{table:best-parkinsons}
\end{table}

>>>>

% Insert here the call for the packages your document requires
\usepackage{graphicx}
\usepackage{tikz}     % For adding axes to plots
\usetikzlibrary{positioning}
\usepackage{graphicx} % For including PDFs
\usepackage{amsmath}  % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs} % For formal tables
\usepackage{siunitx}  % For table aligned by decimal point
\usepackage[colorlinks,allcolors=blue]{hyperref}

@@

@@

\begin{abstract}
A rapid increase in the quantity of data available is allowing all
fields of science to generate more accurate models of multivariate
phenomena. Regression and interpolation become challenging when the
dimension of data is large, especially while maintaining tractable
computational complexity. While regression is a popular approach to
solving approximation problems with high dimension, there are often
advantages to interpolation. This paper presents error bounds for
interpolants on moderately high dimensional problems and contrasts
their performance with that of popular regression techniques.
Empirical results demonstrate the viability of interpolants for high
dimensional approximation problems, suggesting that these techniques
are capable of effectively modeling multivariate phenomena while
maintaining flexibility in different application domains.

@@

\color{red} This section presents the theoretical results bounding the error of
simplicial interpolation. The error analysis relies on simplicial
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straight forward extensions of this error
bound. \color{black}

@@

In summary, the approximation error of a simplicial interpolant tends
quadratically towards zero when approaching observed data only when
the diameter of the simplex is also reduced at a proportional rate.
Only linear convergence to the true function can be achieved in
practice, without the incorporation of additional observations. Notice
that the approximation error is largely determined by the spacing of
observed data. Predictions made by simplices whose vertices are not
well-spaced (i.e. have large diameter, or are nearly contained in a
hyperplane) have higher error. In light of this error bound, an
empirical evaluation of presented algorithms follows.

%% ===================================================================
\section{Data and Empirical Analysis}
\label{sec:data}

The theoretical results constructed in Section \ref{sec:theory} for
simplicial interpolation are promising, however they are difficult to
interpret in context with approximation algorithms that do not have
similar known uniform error bounds. For that reason, this paper
utilizes five different data sets of varying dimension and application
to construct approximations and compare the accuracy of different
techniques.

@@

All approximations are constructed using $k$-fold cross validation as
described in \cite{kohavi1995study} with $k=10$. This approach
randomly partitions data into $k$ nearly equal sized sets. Each
algorithm is then evaluated by constructing an approximation over each
unique union of $k-1$ elements of the partition, making predictions
for points in the remaining element. As a result, each observed data
point is used in the construction of $k-1$ different approximations
and is approximated exactly once. The $k$-fold cross validation method
is data-efficient and provides an unbiased estimate of the expected
prediction error \cite{kohavi1995study}, however it should be noted
that neither this method nor others can provide a universally unbiased
estimator for the variance of prediction error \cite{bengio2004no}.

@@

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c}
    \hline
    Algorithm & Avg. \% Best & Avg. Fit / Prep. Time (s) & Avg. App. Time (s)\\
    \hline
    MARS & $4.6$ & $22.3$ & $1.37 \times 10^{-3}$\\
    SVR & $\mathit{21.1}$ & $0.503$ & $\mathit{1.37 \times 10^{-4}}$\\
    MLP & $\mathbf{42.1}$ & $210.0$ & $1.02 \times 10^{-3}$\\
    Delaunay & $6.0$ & $\mathbf{1.43 \times 10^{-4}}$ & $3.03$\\
    ShepMod & $18.4$ & $0.711$ & $1.49 \times 10^{-4}$\\
    LSHEP & $8.4$ & $1.69$ & $\mathbf{1.16 \times 10^{-4}}$\\
    Voronoi & $0.5$ & $\mathit{0.127}$ & $0.0375$\\
    BoxSplines & $0.2$ & $0.816$ & $5.26 \times 10^{-4}$\\
    \hline
  \end{tabular}
  \caption{This average of Appendix Tables
    \ref{table:best-forest-fire}, \ref{table:best-parkinsons},
    \ref{table:best-weather}, and \ref{table:best-credit-card}
    provides a gross summary of overall results. The columns display
    (weighted equally by data set, \textit{not} points) the average
    frequency with which any algorithm provided the lowest absolute
    error approximation, the average time to fit / prepare, and the
    average time required to approximate one point. Interpolants
    provide the lowest error approximation one third of the time
    across all data, while regressors occupy the other two
    thirds. This result is obtained without any customized tuning or
    preprocessing to maximize the performance of any given
    algorithm. In practice, tuning and preprocessing may have large
    effects on approximation performance.}
  \label{table:avg-performance}
\end{table}

@@

The major advantages to interpolation lie in the near absence of
\textit{fit} time. Delaunay has no preparations required, LSHEP and
ShepMod require pairwise distance calculations to determine the radii
of influence for points in each model. At least hundreds, and
sometimes hundreds of thousands of predictions can be made by the
interpolants before the most widely used regressor (the MLP) finishes
fitting these relatively small data sets. However, the computational
complexities of all interpolants presented are greater than linear in
either dimension or number of points, meaning large enough data (in
dimension and number of observations) would cause interpolation to be
slower.

The theoretical results presented in Section \ref{sec:theory} directly
apply to Delaunay interpolation, however the performance of Delaunay
does not appear significantly better than other algorithms on these
data sets. This observation may be due to the stochastic nature of the
data, but it also speaks to the alignment of the geometry of the
underlying functions with the approximations generated by the
different interpolation methods. The strong performance of other
interpolants suggests that theoretical results similar to those
presented in this work can be achieved for the other interpolants
under reasonable assumptions.

Finally, most of the interpolants presented in this work benefit from
the ability to model any function over a real vector space with a
range that is closed under convex combinations. The results of the
distribution prediction case study indicate that interpolants can
effectively predict CDFs. The error analysis for that work relies on
the KS statistic, which captures the worst part of any prediction and
hence provides a conservatively large estimate of approximation
error. The average absolute errors in the predicted CDFs are always
lower than the KS statistics. However, the KS statistic was chosen
because of the important statistical theory surrounding it as an error
measure. Considering this circumstance, a nonnegligible volume of
predictions provide impressively low levels of error in that final
case study.

@@

This work presents new theoretical results for interpolation in
arbitrary dimension with simplicial approximations. Coinciding with
the theoretical results, an empirical evaluation across real-world
problems demonstrates that interpolants produce competitively accurate
models of multivariate phenomenon when compared with common regressors
for sparse, moderately high dimensional problems.

The various interpolants discussed in this paper have been
demonstrated to effectively approximate multivariate phenomena up to
$30$ dimensions. The underlying constructions are theoretically
straightforward, yet powerful and flexible. Most of the interpolants'
computational complexities make them particularly suitable for
applications in even higher dimension. The major benefits of
interpolation are seen when only a small number of approximations
($\leq 1000$) are made from data and when there are relatively few
data points for the dimension (for empirical results presented,
$\log_d n \leq 5$).

@@

\begin{appendix}

\section*{Appendix}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $0.00984$ & $3.11$ & $7.01$ & $\mathit{11.7}$ & $1090.0$\\
    SVR & $\mathit{0.00402}$ & $\mathbf{0.243}$ & $\mathbf{0.416}$ & $\mathbf{6.19}$ & $1090.0$\\
    MLP & $0.0426$ & $2.63$ & $\mathit{5.27}$ & $14.0$ & $1090.0$\\
    Delaunay & $\mathbf{0.00}$ & $1.98$ & $5.37$ & $13.1$ & $\mathit{1080.0}$\\
    ShepMod & $\mathbf{0.00}$ & $\mathit{1.93}$ & $6.27$ & $16.0$ & $1090.0$\\
    LSHEP & $0.0400$ & $2.17$ & $8.87$ & $19.1$ & $\mathbf{1070.0}$\\
    Voronoi & $0.00982$ & $3.65$ & $7.56$ & $15.6$ & $1090.0$\\
    BoxSpline & $\mathbf{0.00}$ & $2.27$ & $5.91$ & $16.4$ & $1090.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-forest-fire}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit / Prep. Time (s) & App. Time (s) & Total App. Time (s)\\
    \cline{1-2}\cline{4-6}
    MARS & $\mathit{7.7}$ &  & $29.1$ & $1.37 \times 10^{-3}$ & $0.0686$\\
    SVR & $\mathbf{80.2}$ &  & $5.84 \times 10^{-3}$ & $\mathit{6.20 \times 10^{-5}}$ & $\mathit{3.10 \times 10^{-3}}$\\
    MLP & $0.0$ &  & $32.8$ & $8.71 \times 10^{-4}$ & $0.0436$\\
    Delaunay & $3.5$ &  & $\mathbf{2.49 \times 10^{-5}}$ & $0.0385$ & $1.93$\\
    ShepMod & $3.7$ &  & $6.34 \times 10^{-3}$ & $6.44 \times 10^{-5}$ & $3.22 \times 10^{-3}$\\
    LSHEP & $5.1$ &  & $0.0275$ & $\mathbf{4.63 \times 10^{-5}}$ & $\mathbf{2.31 \times 10^{-3}}$\\
    Voronoi & $0.0$ &  & $\mathit{1.82 \times 10^{-4}}$ & $3.96 \times 10^{-4}$ & $0.0198$\\
    BoxSpline & $0.4$ &  & $7.24 \times 10^{-3}$ & $9.78 \times 10^{-5}$ & $4.89 \times 10^{-3}$\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in Table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.}
  \label{table:best-forest-fire}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $0.00948$ & $9.98$ & $20.4$ & $32.9$ & $863.0$\\
    SVR & $0.00138$ & $3.17$ & $7.31$ & $12.6$ & $\mathbf{27.5}$\\
    MLP & $2.39 \times 10^{-5}$ & $\mathit{0.533}$ & $\mathit{1.25}$ & $\mathbf{2.84}$ & $39.3$\\
    Delaunay & $\mathit{3.72 \times 10^{-12}}$ & $1.20$ & $3.50$ & $7.67$ & $30.7$\\
    ShepMod & $\mathbf{0.00}$ & $\mathbf{0.255}$ & $\mathbf{0.908}$ & $\mathit{3.43}$ & $34.5$\\
    LSHEP & $0.00254$ & $2.93$ & $7.16$ & $13.1$ & $\mathit{29.0}$\\
    Voronoi & $\mathbf{0.00}$ & $1.29$ & $3.52$ & $6.87$ & $30.1$\\
    BoxSpline & $0.00287$ & $4.39$ & $9.10$ & $14.8$ & $41.3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit / Prep. Time (s) & App. Time (s) & Total App. Time (s)\\
    \cline{1-2}\cline{4-6}
    MARS & $0.0$ &  & $37.9$ & $2.53 \times 10^{-3}$ & $1.48$\\
    SVR & $0.1$ &  & $0.859$ & $\mathit{1.81 \times 10^{-4}}$ & $\mathit{0.106}$\\
    MLP & $\mathit{32.0}$ &  & $348.0$ & $1.11 \times 10^{-3}$ & $0.653$\\
    Delaunay & $0.0$ &  & $\mathbf{1.90 \times 10^{-4}}$ & $3.69$ & $2.17 \times 10^{3}$\\
    ShepMod & $\mathbf{66.4}$ &  & $1.13$ & $1.82 \times 10^{-4}$ & $0.107$\\
    LSHEP & $0.0$ &  & $2.39$ & $\mathbf{1.44 \times 10^{-4}}$ & $\mathbf{0.0845}$\\
    Voronoi & $1.6$ &  & $\mathit{0.298}$ & $0.0274$ & $16.1$\\
    BoxSpline & $0.0$ &  & $1.26$ & $6.43 \times 10^{-4}$ & $0.377$\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Parkinson's data in Table
    \ref{table:error-parkinsons}. On the right columns are median fit
    time of 5288 points, median time for one approximation, and median
    time approximating 587 points.}
  \label{table:best-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $\mathit{6.45 \times 10^{-15}}$ & $\mathbf{2.70 \times 10^{-14}}$ & $1.66$ & $1.96$ & $\mathit{53.3}$\\
    SVR & $0.0420$ & $0.143$ & $0.148$ & $\mathit{0.860}$ & $119.0$\\
    MLP & $6.89 \times 10^{-5}$ & $0.0337$ & $\mathbf{0.0795}$ & $\mathbf{0.264}$ & $\mathbf{5.31}$\\
    Delaunay & $\mathbf{0.00}$ & $0.187$ & $0.919$ & $2.72$ & $56.3$\\
    ShepMod & $\mathbf{0.00}$ & $0.0957$ & $0.685$ & $2.90$ & $73.2$\\
    LSHEP & $\mathbf{0.00}$ & $\mathit{0.0153}$ & $\mathit{0.106}$ & $1.17$ & $113.0$\\
    Voronoi & $\mathbf{0.00}$ & $0.430$ & $1.28$ & $2.94$ & $83.8$\\
    BoxSpline & $\mathbf{0.00}$ & $0.789$ & $2.06$ & $5.25$ & $119.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-weather}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-weather}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit / Prep. Time (s) & App. Time (s) & Total App. Time (s)\\
    \cline{1-2}\cline{4-6}
    MARS & $10.7$ &  & $0.151$ & $1.17 \times 10^{-4}$ & $0.0304$\\
    SVR & $4.1$ &  & $0.133$ & $\mathbf{9.47 \times 10^{-5}}$ & $\mathbf{0.0246}$\\
    MLP & $\mathbf{56.8}$ &  & $169.0$ & $1.37 \times 10^{-3}$ & $0.356$\\
    Delaunay & $0.1$ &  & $\mathbf{1.16 \times 10^{-4}}$ & $1.55$ & $403.0$\\
    ShepMod & $3.5$ &  & $0.265$ & $1.28 \times 10^{-4}$ & $0.0333$\\
    LSHEP & $\mathit{28.4}$ &  & $0.874$ & $\mathit{9.75 \times 10^{-5}}$ & $\mathit{0.0254}$\\
    Voronoi & $0.2$ &  & $\mathit{0.0112}$ & $0.0270$ & $7.01$\\
    BoxSpline & $0.3$ &  & $0.330$ & $4.06 \times 10^{-4}$ & $0.106$\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Sydney rainfall data in Table
    \ref{table:error-weather}. On the right columns are median fit
    time of 2349 points, median time for one approximation, and median
    time approximating 260 points.}
  \label{table:best-weather}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $5.36$ & $3610.0$ & $4580.0$ & $5450.0$ & $13400.0$\\
    SVR & $0.0138$ & $8.47$ & $15.4$ & $41.9$ & $7700.0$\\
    MLP & $0.00151$ & $\mathbf{1.60}$ & $\mathbf{3.86}$ & $\mathbf{8.34}$ & $\mathbf{604.0}$\\
    Delaunay & $\mathbf{0.00}$ & $\mathit{2.69}$ & $\mathit{13.8}$ & $\mathit{35.0}$ & $\mathit{4840.0}$\\
    ShepMod & $\mathbf{0.00}$ & $4.21$ & $17.3$ & $51.6$ & $6510.0$\\
    LSHEP & $1.27$ & $199.0$ & $260.0$ & $343.0$ & $6530.0$\\
    Voronoi & $\mathit{2.89 \times 10^{-10}}$ & $14.7$ & $32.8$ & $52.9$ & $4860.0$\\
    BoxSpline & $0.00740$ & $20.8$ & $40.9$ & $79.6$ & $7640.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-credit-card}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-credit-card}
\end{table}
